---
title: "SpeedDating market segmentation"
author: "T. Evgeniou; adapted by team"
output:
  html_document:
    css: styles/default.css
    theme: paper
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    includes:
      in_header: styles/default.sty
always_allow_html: yes
---

> **IMPORTANT**: Please make sure you create a copy of this file with a customized name, so that your work (e.g. answers to the questions) is not over-written when you pull the latest content from the course github. 
This is a **template process for market segmentation based on survey data**, using the  [Boats cases A](http://inseaddataanalytics.github.io/INSEADAnalytics/Boats-A-prerelease.pdf) and  [B](http://inseaddataanalytics.github.io/INSEADAnalytics/Boats-B-prerelease.pdf).

All material and code is available at the [INSEAD Data Analytics for Business](http://inseaddataanalytics.github.io/INSEADAnalytics/) website and github. Before starting, make sure you have pulled the [course files](https://github.com/InseadDataAnalytics/INSEADAnalytics) on your github repository.  As always, you can use the `help` command in Rstudio to find out about any R function (e.g. type `help(list.files)` to learn what the R function `list.files` does).


<hr>\clearpage

# The Business Questions

This process can be used as a (starting) template for projects like the one described in the [Boats cases A](http://inseaddataanalytics.github.io/INSEADAnalytics/Boats-A-prerelease.pdf) and  [B](http://inseaddataanalytics.github.io/INSEADAnalytics/Boats-B-prerelease.pdf). For example (but not only), in this case some of the business questions were: 

- What are the main purchase drivers of the customers (and prospects) of this company? 

- Are there different market segments? Which ones? Do the purchase drivers differ across segments? 

- What (possibly market segment specific) product development or brand positioning strategy should the company follow in order to increase its sales? 

See for example some of the analysis of this case in  these slides: <a href="http://inseaddataanalytics.github.io/INSEADAnalytics/Sessions2_3 Handouts.pdf"  target="_blank"> part 1</a> and <a href="http://inseaddataanalytics.github.io/INSEADAnalytics/Sessions4_5 Handouts.pdf"  target="_blank"> part 2</a>.

<hr>\clearpage

# The Process

The "high level" process template is split in 3 parts, corresponding to the course sessions 3-4, 5-6, and an optional last part: 

1. *Part 1*: We use some of the survey questions (e.g. in this case the first 29 "attitude" questions) to find **key customer descriptors** ("factors") using *dimensionality reduction* techniques described in the [Dimensionality Reduction](http://inseaddataanalytics.github.io/INSEADAnalytics/CourseSessions/Sessions23/FactorAnalysisReading.html) reading of Sessions 3-4.

2. *Part 2*: We use the selected customer descriptors to **segment the market** using *cluster analysis* techniques described in the [Cluster Analysis ](http://inseaddataanalytics.github.io/INSEADAnalytics/CourseSessions/Sessions45/ClusterAnalysisReading.html) reading of Sessions 5-6.

3. *Part 3*: For the market segments we create, we will use *classification analysis* to classify people based on whether or not they have purchased a product and find what are the **key purchase drivers per segment**. For this part we will use [classification analysis ](http://inseaddataanalytics.github.io/INSEADAnalytics/CourseSessions/Sessions67/ClassificationAnalysisReading.html) techniques.

Finally, we will use the results of this analysis to make business decisions e.g. about brand positioning, product development, etc depending on our market segments and key purchase drivers we find at the end of this process.


```{r setuplibraries, echo=FALSE, message=FALSE}
suppressWarnings(source("lib/library.R"))
# Package options
suppressWarnings(ggthemr('fresh'))  # ggplot theme
opts_knit$set(progress=FALSE, verbose=FALSE)
opts_chunk$set(echo=FALSE, fig.align="center", fig.width=10, fig.height=6.35, results="asis")
options(knitr.kable.NA = '')
```

<hr>\clearpage

# The Data

First we load the data to use (see the raw .Rmd file to change the data file as needed):

```{r setupdata1E, echo=TRUE, tidy=TRUE}
# Please ENTER the name of the file with the data used. The file should be a .csv with one row per observation (e.g. person) and one column per attribute. Do not add .csv at the end, make sure the data are numeric.
datafile_name = "data/SpeedDatingData_clean.csv"

# Please enter the minimum number below which you would like not to print - this makes the readability of the tables easier. Default values are either 10e6 (to print everything) or 0.5. Try both to see the difference.
MIN_VALUE = 0.5

# Please enter the maximum number of observations to show in the report and slides. 
# DEFAULT is 10. If the number is large the report may be slow.
max_data_report = 30
```

```{r}
ProjectData <- read.delim(file = datafile_name, sep = ";")
ProjectData <- data.matrix(ProjectData) 
ProjectData_INITIAL <- ProjectData
```

<hr>\clearpage

# Part 1: Key Customer Characteristics

The code used here is along the lines of the code in the session 3-4 reading  [FactorAnalysisReading.Rmd](https://github.com/InseadDataAnalytics/INSEADAnalytics/blob/master/CourseSessions/Sessions23/FactorAnalysisReading.Rmd). We follow the process described in the [Dimensionality Reduction ](http://inseaddataanalytics.github.io/INSEADAnalytics/CourseSessions/Sessions23/FactorAnalysisReading.html) reading. 

In this part we also become familiar with:

1. Some visualization tools;
2. Principal Component Analysis and Factor Analysis;
3. Introduction to machine learning methods;

(All user inputs for this part should be selected in the code chunk in the raw .Rmd file) 

```{r setupfactor, echo=TRUE, tidy=TRUE}
# factor_attributes_used = c("goal", "attr1_1",	"sinc1_1",	"intel1_1", "fun1_1","amb1_1", "shar1_1", "attr1_1",	"sinc2_1",	"intel2_1", "fun2_1","amb2_1", "shar2_1", "sinc3_1",	"intel3_1", "fun3_1","amb3_1", "sinc5_1",	"intel5_1", "fun5_1","amb5_1")

factor_attributes_used = c(10:47)

# Please ENTER the selection criterions for the factors to use. 
# Choices: "eigenvalue", "variance", "manual"
factor_selectionciterion = "manual"

# Please ENTER the desired minumum variance explained 
# (Only used in case "variance" is the factor selection criterion used). 
minimum_variance_explained = 65  # between 1 and 100

# Please ENTER the number of factors to use 
# (Only used in case "manual" is the factor selection criterion used).
manual_numb_factors_used = 10

# Please ENTER the rotation eventually used (e.g. "none", "varimax", "quatimax", "promax", "oblimin", "simplimax", and "cluster" - see help(principal)). Default is "varimax"
rotation_used = "varimax"
```

```{r}
factor_attributes_used <- intersect(factor_attributes_used, 1:ncol(ProjectData))
ProjectDataFactor <- ProjectData[,factor_attributes_used]
ProjectDataFactor <- ProjectData <- data.matrix(ProjectDataFactor)

# remove rows with empty cells
ProjectDataFactor <- ProjectDataFactor[rowSums(is.na(ProjectDataFactor)) == 0,]

```

## Steps 1-2: Check the Data 

Start by some basic visual exploration of, say, a few data:

```{r}
rownames(ProjectDataFactor) <- paste0("Obs.", sprintf("%02i", 1:nrow(ProjectDataFactor)))
iprint.df(t(head(round(ProjectDataFactor, 2), max_data_report)))
```

The data we use here have the following descriptive statistics: 

```{r}
iprint.df(round(my_summary(ProjectDataFactor), 2))
```

## Step 3: Check Correlations

This is the correlation matrix of the customer responses to the `r ncol(ProjectDataFactor)` attitude questions - which are the only questions that we will use for the segmentation (see the case):

```{r}

thecor = round(cor(ProjectDataFactor),2)
iprint.df(round(thecor,2), scale=TRUE)
```

**Questions**

1. Do you see any high correlations between the responses? Do they make sense? 
2. What do these correlations imply?

**Answers:**

*
*
*
*
*
*
*
*
*
*

## Step 4: Choose number of factors

Clearly the survey asked many redundant questions (can you think some reasons why?), so we may be able to actually "group" these 29 attitude questions into only a few "key factors". This not only will simplify the data, but will also greatly facilitate our understanding of the customers.

To do so, we use methods called [Principal Component Analysis](https://en.wikipedia.org/wiki/Principal_component_analysis) and [factor analysis](https://en.wikipedia.org/wiki/Factor_analysis) as also discussed in the [Dimensionality Reduction readings](http://inseaddataanalytics.github.io/INSEADAnalytics/CourseSessions/Sessions23/FactorAnalysisReading.html). We can use two different R commands for this (they make slightly different information easily available as output): the command `principal` (check `help(principal)` from R package [psych](http://personality-project.org/r/psych/)), and the command `PCA` from R package [FactoMineR](http://factominer.free.fr) - there are more packages and commands for this, as these methods are very widely used.  

```{r}
# Here is how the `principal` function is used 
UnRotated_Results<-principal(ProjectDataFactor, nfactors=ncol(ProjectDataFactor), rotate="none",score=TRUE)
UnRotated_Factors<-round(UnRotated_Results$loadings,2)
UnRotated_Factors<-as.data.frame(unclass(UnRotated_Factors))
colnames(UnRotated_Factors)<-paste("Comp",1:ncol(UnRotated_Factors),sep="")
```

```{r}
# Here is how we use the `PCA` function 
Variance_Explained_Table_results<-PCA(ProjectDataFactor, graph=FALSE)
Variance_Explained_Table<-Variance_Explained_Table_results$eig
Variance_Explained_Table_copy<-Variance_Explained_Table

rownames(Variance_Explained_Table) <- paste("Component", 1:nrow(Variance_Explained_Table), sep=" ")
colnames(Variance_Explained_Table) <- c("Eigenvalue", "Pct of explained variance", "Cumulative pct of explained variance")
```

Let's look at the **variance explained** as well as the **eigenvalues** (see session readings):

```{r}
iprint.df(round(Variance_Explained_Table, 2))
```

```{r}
eigenvalues  <- Variance_Explained_Table[, "Eigenvalue"]
df           <- cbind(as.data.frame(eigenvalues), c(1:length(eigenvalues)), rep(1, length(eigenvalues)))
colnames(df) <- c("eigenvalues", "components", "abline")
iplot.df(melt(df, id="components"))
```

**Questions:**

1. Can you explain what this table and the plot are? What do they indicate? What can we learn from these?
2. Why does the plot have this specific shape? Could the plotted line be increasing?
3. What characteristics of these results would we prefer to see? Why?

**Answers**

*
*
*
*
*
*
*
*
*
*

## Step 5: Interpret the factors

Let's now see how the "top factors" look like.

```{r}
if (factor_selectionciterion == "eigenvalue")
  factors_selected = sum(Variance_Explained_Table_copy[,1] >= 1)
if (factor_selectionciterion == "variance")
  factors_selected = 1:head(which(Variance_Explained_Table_copy[,"cumulative percentage of variance"]>= minimum_variance_explained),1)
if (factor_selectionciterion == "manual")
  factors_selected = manual_numb_factors_used
```

To better visualize them, we will use what is called a "rotation". There are many rotations methods. In this case we selected the `r rotation_used` rotation. For our data, the `r factors_selected` selected factors look as follows after this rotation:

```{r}
Rotated_Results<-principal(ProjectDataFactor, nfactors=max(factors_selected), rotate=rotation_used,score=TRUE)
Rotated_Factors<-round(Rotated_Results$loadings,2)
Rotated_Factors<-as.data.frame(unclass(Rotated_Factors))
colnames(Rotated_Factors)<-paste("Comp.",1:ncol(Rotated_Factors),sep="")

sorted_rows <- sort(Rotated_Factors[,1], decreasing = TRUE, index.return = TRUE)$ix
Rotated_Factors <- Rotated_Factors[sorted_rows,]

iprint.df(Rotated_Factors, scale=TRUE)
```

To better visualize and interpret the factors we often "suppress" loadings with small values, e.g. with absolute values smaller than 0.5. In this case our factors look as follows after suppressing the small numbers:

```{r}
Rotated_Factors_thres <- Rotated_Factors
Rotated_Factors_thres[abs(Rotated_Factors_thres) < MIN_VALUE]<-NA
colnames(Rotated_Factors_thres)<- colnames(Rotated_Factors)
rownames(Rotated_Factors_thres)<- rownames(Rotated_Factors)

iprint.df(Rotated_Factors_thres, scale=TRUE)
```

**Questions**

1. What do the first couple of factors mean? Do they make business sense?
2. How many factors should we choose for this data/customer base? Please try a few and explain your final choice based on a) statistical arguments, b) on interpretation arguments, c) on business arguments (**you need to consider all three types of arguments**)
3. How would you interpret the factors you selected?
4. What lessons about data science do you learn when doing this analysis? Please comment.

**Answers**

*
*
*
*
*
*
*
*
*
*

## Step 6:  Save factor scores

We can now either replace all initial variables used in this part with the factors scores or just select one of the initial variables for each of the selected factors in order to represent that factor. Here is how the factor scores  are for the first few respondents:

```{r}
NEW_ProjectData <- round(Rotated_Results$scores[,1:factors_selected,drop=F],2)
colnames(NEW_ProjectData)<-paste("DV (Factor)",1:ncol(NEW_ProjectData),sep=" ")

iprint.df(t(head(NEW_ProjectData, 10)), scale=TRUE)
```

**Questions**

1. Can you describe some of the people using the new derived variables (factor scores)?
2. Which of the 29 initial variables would you select to represent each of the factors you selected?

**Answers**

*
*
*
*
*
*
*
*
*
*



# Part 4: Regression Analysis

```{r}

ProjectDataRegr<-data.frame(ProjectData_INITIAL)

decision.attr1<-lm(dec_o~attr1_1+sinc1_1+intel1_1+fun1_1+amb1_1+shar1_1+attr2_1+sinc2_1+intel2_1+fun2_1+amb2_1+shar2_1+attr3_1+sinc3_1+intel3_1+fun3_1+amb3_1+attr4_1+sinc4_1+intel4_1+fun4_1+amb4_1+shar4_1,ProjectDataRegr)

iprint.df(coef(summary(decision.attr1)), scale=TRUE)

plot(decision.attr1)

```

<hr>\clearpage

