---
title: "SpeedDating - prediciting success in speed dating"
author: "T. Evgeniou; adapted by team"
output:
  html_document:
    css: styles/default.css
    theme: paper
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    includes:
      in_header: styles/default.sty
always_allow_html: yes
---

<hr>\clearpage

# Speed dating insights

## The "Business Decision"

We would like to understand how people rate themselves across **five attributes** (attractiveness, sincerity, intelligence, fun and ambition) and how this compares to the ratings they get from their dating partners.

## The Data

```{r setuplibraries, echo=FALSE, message=FALSE}
suppressWarnings(source("lib/library.R"))
# Package options
suppressWarnings(ggthemr('fresh'))  # ggplot theme
opts_knit$set(progress=FALSE, verbose=FALSE)
opts_chunk$set(echo=FALSE, fig.align="center", fig.width=10, fig.height=6.35, results="asis")
options(knitr.kable.NA = '')

datafile_name = "data/SpeedDatingData_clean.csv"

ProjectData <- read.delim(file = datafile_name, sep = ";")

ProjectData_INITIAL <- ProjectData

max_data_report<-5

independent_variables= c(25:30, 36, 40, 45:67) 

```

We used the data from speed dating experiment conducted at Columbia Business School <a href="https://www.kaggle.com/annavictoria/speed-dating-experiment">available on kaggle</a>. Refer to <a href="https://www.kaggle.com/annavictoria/speed-dating-experiment/downloads/speed-dating-experiment.zip"> docummentation</a href> for more details about used attributes or jump to the <a href="SpeedDatingMarketSegmentation.html">segmentation section</a href> for short description.

This is how the first `r min(max_data_report, nrow(ProjectData))` out of the total of `r nrow(ProjectData)` rows look:

```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
# let's make the data into data.matrix classes so that we can easier visualize them
ProjectDataMatrix = data.matrix(ProjectData)

knitr::kable({
  df <- t(head(round(ProjectDataMatrix[,independent_variables],2), max_data_report))
  colnames(df) <- sprintf("%02d", 1:ncol(df))
  df
})
```

## The Process

The "high level" process template to derive high level insights from the speed dating database is split in 3 parts: 

1. *Part 1*: We use descrictive statistics (e.g. mean, standard deviation) to understand whether men and women tend to overvalue or undervalue themselves across the five studied attributes

2. *Part 2*: We run hypothesis testing to estimate whether we can accept or reject with 99% confidence that men and women overvalue themselves


3. *Part 3*: We plot the self-ratings of men and women against the ratings given by their dating-partners and we look for patterns of over/undervaluation across the two sexes


<hr>\clearpage

# Descriptive Statistics

## Self-perception vs partner's rating

For each participant we estimated the difference between his/her self-rating and that of the dating partner for the total rating as well as for each of the five attributes. Then we computed the mean and st deviation of that difference for the total population and then for men and women separately. Looking at the fraction of people that overvalue or unervalue themselves can provide us with meaningful insights about self-perception and dating behavior

```{r echo=FALSE}
ProjectData$Delta_attr<-ProjectData$attr_o - ProjectData$attr5_1
ProjectData$Delta_sinc<-ProjectData$sinc_o - ProjectData$sinc5_1
ProjectData$Delta_intel<-ProjectData$intel_o - ProjectData$intel5_1
ProjectData$Delta_fun<-ProjectData$fun_o - ProjectData$fun5_1
ProjectData$Delta_amb<-ProjectData$amb_o - ProjectData$amb5_1
ProjectData$Delta_avg<-(ProjectData$Delta_attr+ProjectData$Delta_sinc+ProjectData$Delta_intel+ProjectData$Delta_fun+ProjectData$Delta_amb)/5
ProjectData <- ProjectData[!is.na(ProjectData$Delta_avg),]

insights_row_names<-c("Overall","Men","Women","SD Men","SD Women","Undervalued (Overall)","Overvalued (Overall)","On spot (Overall)","Undervalued (men)","Overvalued (men)","On spot (men)","Undervalued (women)","Overvalued (women)","On spot (women)")

insights_col_names<-c("Overall_Avg_delta","%","Avg_Attr_Delta","Attr_%","Avg_Sinc_Delta","Sinc_%","Avg_Intel_Delta","Intel_%","Avg_Fun_Delta","Fun_%","Avg_Amb_Delta","Amb_%")

ProjectData_male<-subset(ProjectData,ProjectData$gender==1)
ProjectData_female<-subset(ProjectData,ProjectData$gender==0)

Project_insights<-NULL

for (i in c(201,196,197,198,199,200)) {

insights_1<-c(
round(mean(ProjectData[,i]),2),
round(mean(ProjectData_male[,i]),2),
round(mean(ProjectData_female[,i]),2),
round(sd(ProjectData_male[,i]),2),
round(sd(ProjectData_female[,i]),2),
length(ProjectData[ProjectData[,i]<0,1]),
length(ProjectData[ProjectData[,i]>0,1]),
length(ProjectData[ProjectData[,i]==0,1]),
length(ProjectData_male[ProjectData_male[,i]<0,1]),
length(ProjectData_male[ProjectData_male[,i]>0,1]),
length(ProjectData_male[ProjectData_male[,i]==0,1]),
length(ProjectData_female[ProjectData_female[,i]<0,1]),
length(ProjectData_female[ProjectData_female[,i]>0,1]),
length(ProjectData_female[ProjectData_female[,i]==0,1])
)

insights_2<-c(
"","","","","",
paste(percent(length(ProjectData[ProjectData[,i]<0,1])/length(ProjectData[,1]))),
paste(percent(length(ProjectData[ProjectData[,i]>0,1])/length(ProjectData[,1]))),
paste(percent(length(ProjectData[ProjectData[,i]==0,1])/length(ProjectData[,1]))),
paste(percent(length(ProjectData_male[ProjectData_male[,i]<0,1])/length(ProjectData_male[,1]))),
paste(percent(length(ProjectData_male[ProjectData_male[,i]>0,1])/length(ProjectData_male[,1]))),
paste(percent(length(ProjectData_male[ProjectData_male[,i]==0,1])/length(ProjectData_male[,1]))),
paste(percent(length(ProjectData_female[ProjectData_female[,i]<0,1])/length(ProjectData_female[,1]))),
paste(percent(length(ProjectData_female[ProjectData_female[,i]>0,1])/length(ProjectData_female[,1]))),
paste(percent(length(ProjectData_female[ProjectData_female[,i]==0,1])/length(ProjectData_female[,1])))
)

insights<-cbind(insights_1,insights_2)
Project_insights<-cbind(Project_insights,insights)
}

dimnames(Project_insights)<-list(insights_row_names,insights_col_names)

iprint.df(Project_insights, scale=TRUE)

```

## Importance of each attribute

Next, we estimate the average score that men and women get from their dating partners' and the importance of its attribute for mens' and womens' matching decision.

```{r echo=FALSE}

Project_insights_2<-cbind(
c(round(mean(ProjectData_male$attr_o),2),
round(mean(ProjectData_female$attr_o),2),
paste(percent(mean(ProjectData_male$attr1_1)/100)),
paste(percent(mean(ProjectData_female$attr1_1)/100))
),
c(round(mean(ProjectData_male$sinc_o),2),
round(mean(ProjectData_female$sinc_o),2),
paste(percent(mean(ProjectData_male$sinc1_1)/100)),
paste(percent(mean(ProjectData_female$sinc1_1)/100))
),
c(round(mean(ProjectData_male$intel_o),2),
round(mean(ProjectData_female$intel_o),2),
paste(percent(mean(ProjectData_male$intel1_1)/100)),
paste(percent(mean(ProjectData_female$intel1_1)/100))
),
c(round(mean(ProjectData_male$fun_o),2),
round(mean(ProjectData_female$fun_o),2),
paste(percent(mean(ProjectData_male$fun1_1)/100)),
paste(percent(mean(ProjectData_female$fun1_1)/100))
),
c(round(mean(ProjectData_male$amb_o),2),
round(mean(ProjectData_female$amb_o),2),
paste(percent(mean(ProjectData_male$amb1_1)/100)),
paste(percent(mean(ProjectData_female$amb1_1)/100))
)
)


dimnames(Project_insights_2)<-list(c("Average Men's Score","Average Women's Score","Importance for Men","Importance for Women"),c("Attractiveness","Sincerity","Intelligence","Fun","Ambition"))

iprint.df(Project_insights_2, scale=TRUE)

```

## Discussion

Based on the above statistics, we can derive the following insights:

* Both, men and women rate themselves on average better than their dating partners
* There is no significant difference in the gap between self- and external perception between men and women, the only remarkable difference is regarding sincerity. Women think of themselves way more sincere than they are perceived by their dates – this effect is less distinct vice versa
* Attractiveness is the by far most important factor at dates for men. 
* Women seem to be overall more balanced in their preferences with attractiveness, sincerity, intelligence and fun being almost equally rated.
* There are no major differences between average ratings for men and women on average
* Attractiveness tends to be the lowest-rated attribute for both genders although / or because of it is one of the attributes with the highest importance
* Intelligence and Sincerity are highest-rated (would be interesting to analyze if those factors have a high ipact on match, I guess not)
* Overall, there is no significant difference between male and female valuations during speed-dating, neither in self-perception nor in rating the partners. However, it is yet to be analyzed how / if their ratings have differing influences on final decisions

# Hypothesis Testing

To validate the significance of the above inisghts, we run a one-sided hypothesis testing

Hypotheses:

*H0: mu-delta >= 0*

*HA: mu-delta <  0  (meaning that self-rating is higher than partner's rating, i.e. overrating)*

```{r echo=FALSE}

hypothesis_row_names<-c("Mean (men)","SD (men)","Sample size (men)","t-value (men)","Critical Value at 1% (men)","Statistical Decision (men)","","Mean (women)","SD (women)","Sample size (women)","t-value (women)","Critical Value at 1% (women)","Statistical Decision (women)")

hypothesis_col_names<-c("Attractiveness","Sincerity","Intelligence","Fun","Ambition")

Hypothesis_testing<-NULL

for (i in c(196,197,198,199,200)) {

average_men<-mean(ProjectData_male[,i])
sd_men<-sd(ProjectData_male[,i])
size_men<-length(ProjectData_male[,i])
tvalue_men<-average_men/(sqrt(sd_men/size_men))
critical_value_men<-qt(0.05,size_men-1)

average_women<-mean(ProjectData_female[,i])
sd_women<-sd(ProjectData_female[,i])
size_women<-length(ProjectData_female[,i])
tvalue_women<-average_women/(sqrt(sd_women/size_women))
critical_value_women<-qt(0.01,size_women-1)

Hypothesis_testing<-cbind(Hypothesis_testing,c(
round(average_men,2),
round(sd_men,2),
size_men,
round(tvalue_men,3),
round(critical_value_men,3),
if(tvalue_men<critical_value_men) {"Reject"} else {"Accept"},
"",
round(average_women,2),
round(sd_women,2),
size_women,
round(tvalue_women,3),
round(critical_value_women,3),
if(tvalue_women<critical_value_women) {"Reject"} else {"Accept"}
))}

dimnames(Hypothesis_testing)<-list(hypothesis_row_names,hypothesis_col_names)

iprint.df(Hypothesis_testing, scale=TRUE)

```

On the basis of these results, we conclude that men and women tend to overrate themselves across all five attributes.

<hr>\clearpage

# Plots

##Attractiveness

```{r echo=FALSE}

par(mfrow = c(1, 2))

plot(tapply(ProjectData_male$attr_o,factor(ProjectData_male$attr5_1),mean),xlim=c(0,10),ylim=c(0,10),main="Men", xlab="Self-perception", ylab="Others perception",col = 'blue',pch = 16)
plot(tapply(ProjectData_female$attr_o,factor(ProjectData_female$attr5_1),mean),xlim=c(0,10),ylim=c(0,10), main="Women", xlab="Self-perception", ylab="",col = 'red',pch = 16)

```

* Men show a high-degree of self-awareness regarding their attractiveness relative to other men.
* Men tend to consider the women they dated fairly attractive regardless of how these women are perceiving themselves. Interestingly though, women who consider themselves as the most attracive (score: 8), are perceived as the least attractive by their dating partners.


##Sincerity

```{r echo=FALSE}

par(mfrow = c(1, 2))

plot(tapply(ProjectData_male$sinc_o,factor(ProjectData_male$sinc5_1),mean),xlim=c(0,10),ylim=c(0,10),main="Men", xlab="Self-perception", ylab="Others perception",col = 'blue',pch = 16)
plot(tapply(ProjectData_female$sinc_o,factor(ProjectData_female$sinc5_1),mean),xlim=c(0,10),ylim=c(0,10), main="Women", xlab="Self-perception", ylab="",col = 'red',pch = 16)

```

* Both men and women are perceived by their dating partners as fairly sincere (score: 7+), regardless of their self-perception.


##Intelligence

```{r echo=FALSE}

par(mfrow = c(1, 2))

plot(tapply(ProjectData_male$intel_o,factor(ProjectData_male$intel5_1),mean),xlim=c(0,10),ylim=c(0,10),main="Men", xlab="Self-perception", ylab="Others perception",col = 'blue',pch = 16)
plot(tapply(ProjectData_female$intel_o,factor(ProjectData_female$intel5_1),mean),xlim=c(0,10),ylim=c(0,10), main="Women", xlab="Self-perception", ylab="",col = 'red',pch = 16)

```

* Men who rate themselves very low with regards to their intelligence, are considered on average as the most intelligent by their dating partners, indicating that very smart people might also be the most modest.
* Women receive on average the same rating from their partners, regardless of how they pereceive themselves.

##Fun

```{r echo=FALSE}

par(mfrow = c(1, 2))

plot(tapply(ProjectData_male$fun_o,factor(ProjectData_male$fun5_1),mean),xlim=c(0,10),ylim=c(0,10),main="Men", xlab="Self-perception", ylab="Others perception",col = 'blue',pch = 16)
plot(tapply(ProjectData_female$fun_o,factor(ProjectData_female$fun5_1),mean),xlim=c(0,10),ylim=c(0,10), main="Women", xlab="Self-perception", ylab="",col = 'red',pch = 16)

```

* Men show a high-degree of self-awareness with regards to how funny they are relative to other men.


##Ambition

```{r echo=FALSE}

par(mfrow = c(1, 2))

plot(tapply(ProjectData_male$amb_o,factor(ProjectData_male$amb5_1),mean),xlim=c(0,10),ylim=c(0,10),main="Men", xlab="Self-perception", ylab="Others perception",col = 'blue',pch = 16)
plot(tapply(ProjectData_female$amb_o,factor(ProjectData_female$amb5_1),mean),xlim=c(0,10),ylim=c(0,10), main="Women", xlab="Self-perception", ylab="",col = 'red',pch = 16)

```

* Men with very low ambitions are easily recognized as such by their dating partners.
* Both men and women receive on average the same rating from their partners (score: ~7), regardless of how they pereceive themselves.

<hr>\clearpage
